VM-Centric Snapshot Deduplication for Converged Cloud Architectures

### 2014-07-18 snapshot_system/paper_background.tex

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so a number of techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%to reduce memory requirements at the cost of some loss of deduplication
%efficiency.
Additional inline deduplication and  approximation techniques
are studied in ~\cite{extreme_binning09,sparseindex09,Srinivasan2012,WeiZhangIEEE}.  
%and a parallel batch solution for cluster-based deduplication is 
%studied in ~\cite{WeiZhangIEEE}. 
\comments{
All of the above approaches have focused on optimization of deduplication
efficiency or speed, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
}
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

>>>>

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so a number of techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses an in-memory Bloom filter to identify non-duplicates and a prefetching cache for potential
duplicates that might hit in the future.
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
%Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
%to reduce memory requirements at the cost of some loss of deduplication
%efficiency.
Additional inline deduplication and  approximation techniques
are studied in ~\cite{extreme_binning09,sparseindex09,Srinivasan2012,WeiZhangIEEE}.  
%and a parallel batch solution for cluster-based deduplication is 
%studied in ~\cite{WeiZhangIEEE}. 
\comments{
All of the above approaches have focused on optimization of deduplication
efficiency or speed, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
}
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.

### 2014-07-18 snapshot_system/paper_del.tex

\begin{figure}[htbp]
  \centering
  \epsfig{file=images/deletion.pdf, width=3.25in}
  \caption{Fast approximate deletion}
  \label{fig:deletion_flow}
\end{figure}

@@

%{\bf Discussion}
We estimate storage leakage size and how often leak repair should be conducted.
Assume that  a VM keeps $h$ snapshots in backup storage, creates and deletes one snapshot
every day. Let $u$ be the number of chunks brought by initial backup for a VM, $\Delta u$ be the average
number of additional chunks added from one version to next snapshot version. Then the total number of 
chunks in a VM's snapshot store is about:
$
U = u + (h-1)\Delta u.
$

@@

For each snapshot deletion, the number of chunks to be deleted is nearly identical to the number of
newly added chunks $\Delta u$. 
Let $R$ be the total number of runs of approximate deletion between two consecutive 
repairs. We estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \epsilon \Delta u.
\]

When leakage ratio $L/U$ exceeds a pre-defined threshold $\tau$, we trigger a leak repair. Namely,

>>>>

\begin{figure}[htbp]
\vspace{2em}
  \centering
  \epsfig{file=images/deletion.pdf, width=3.35in}
  \caption{Fast approximate deletion}
  \label{fig:deletion_flow}
\end{figure}

@@

%{\bf Discussion}
We estimate storage leakage size and how often leak repair should be conducted.
Assume that  a VM keeps $h$ snapshots in backup storage, creates and deletes one snapshot
every day. Let $u$ be the number of chunks brought by initial backup for a VM, $\Delta u$ be the average
number of additional chunks added from one version to next snapshot version. Then the total number of 
chunks in a VM's snapshot store is about:
\[
U = u + (h-1)\Delta u.
\]

@@

For each snapshot deletion, the number of chunks to be deleted is nearly identical to the number of
newly added chunks $\Delta u$. 
Let $R$ be the total number of runs of approximate deletion between two consecutive 
repairs. We estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \epsilon \Delta u.
\]
When leakage ratio $L/U$ exceeds a pre-defined threshold $\tau$, we trigger a leak repair. Namely,

### 2014-07-18 snapshot_system/paper_eval.tex

Table~\ref{tab:leakage} shows the average accumulated storage leakage in terms of percentage of
storage space per VM caused  by approximate deletions.
In this experiment, we select 105 VMs and let all the VMs accumulate 10 snapshots, 
then start to delete those snapshots one by one in reverse order.
As we know the actual storage needs after each snapshot creation,
the storage leakage can be detected by
comparing the size of remaining data in use after deletion to the correct number.
Row 1 in  Table~\ref{tab:leakage} 
is the deletion step. 3 means that version 3 is deleted. 
Row  2 is the predicted leakage using Formula~\ref{eq:leakrepair} from Section~\ref{sect:delete}
given $\Delta u/u=0.025$,
while Row 3 lists the actual average leakage measured during the experiment for all the VMs. 
The Bloom filter setting is based on $\Delta u/u=0.025$.
After 9 snapshot deletions, the actual leakage ratio reaches 0.0015 and this means that
there is only 1.5MB space leaked for every 1GB of stored data.
The actual leakage can reach  4.1\% after  245 deletions and such a repair is needed every 12 days
to the  leakage under 5\%. 
This experiment shows that the leakage of our approximate snapshot deletion is very small, 
below the estimated number.

The space cost for each snapshot deletion is insignificant.
Leakage repair for each VM needs less than 85MB of memory
considering each reference consumes 8 bytes plus  1 mark bit and  
each VM snapshot has 40GB backup data with about 10 million chunks. 
This VM-specific leakage  repair takes  less than half an hour for each VM. 
%when disk bandwidth is controlled to 50MB/s. 
%When using the traditional mark-and-sweep~\cite{Guo2011,Fabiano2013} in VO for 2500 VMs, 
%a total of 212GB memory
%space on 100 node is accessed for processing each VM and each machine requires 2.12GB.
%To control the garbage growth, such a process needs to be conducted much more frequently than our
%VM-centric leakage repair.

>>>>

Table~\ref{tab:leakage} shows the average accumulated storage leakage in terms of percentage of
storage space per VM caused  by approximate deletions.
In this experiment, we select 105 VMs and let all the VMs accumulate 10 snapshots, 
then start to delete those snapshots one by one in reverse order.
As we know the actual storage needs after each snapshot creation,
the storage leakage can be detected by
comparing the size of remaining data in use after deletion to the correct number.
Row 1 in  Table~\ref{tab:leakage} 
is the deletion step. Entry value 3 in this row   means that Version 3 is deleted. 
Row  2 is the predicted leakage using Formula~\ref{eq:leakrepair} from Section~\ref{sect:delete}
given $\Delta u/u=0.025$,
while Row 3 lists the actual average leakage measured during the experiment for all the VMs. 
The Bloom filter setting is based on $\Delta u/u=0.025$.
After 9 snapshot deletions, the actual leakage ratio reaches 0.0015 and this means that
there is only 1.5MB space leaked for every 1GB of stored data.
The actual leakage can reach  4.1\% after  245 deletions.
% and such a repair is needed every 12 days to the  leakage under 5\%. 
This experiment shows that the leakage of our approximate snapshot deletion is very small, 
below the estimated number.

\comments{
The space cost for each snapshot deletion is insignificant.
Leakage repair for each VM needs less than 85MB of memory
considering each reference consumes 8 bytes plus  1 mark bit and  
each VM snapshot has 40GB backup data with about 10 million chunks. 
This VM-specific leakage  repair takes  less than half an hour for each VM. 
}
%when disk bandwidth is controlled to 50MB/s. 
%When using the traditional mark-and-sweep~\cite{Guo2011,Fabiano2013} in VO for 2500 VMs, 
%a total of 212GB memory
%space on 100 node is accessed for processing each VM and each machine requires 2.12GB.
%To control the garbage growth, such a process needs to be conducted much more frequently than our
%VM-centric leakage repair.
